---
title: "Biomarker Predictivity"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(parallel)
library(foreach)
library(broom)
library(caret)
```

## Independent Validation on METABRIC Data

Here, I validate the pre-processing and selection method used by showing that the biomarkers selected on TCGA BRCA are useful for classification on METABRIC data


### Custom Functions

```{r}
# To perform ANOVA
run_anova <- function(train_data, train_labels, num_cpus=40) {
  num_cpus <- 40
  cl <- makeCluster(num_cpus)
  doParallel::registerDoParallel(cl)
  anova_summary <- foreach(i = 2:ncol(train_data)) %dopar% {
    column <- names(train_data[i]) 
    avz <- broom:: tidy(aov(train_data[,i][[1]] ~ unlist(train_labels$cancer_subtype))) # Each ANOVA test iterating through each column
    return(c(gene=column, f_value=avz$statistic[[1]], p_value=avz$p.value[[1]]))
  }
  stopCluster(cl)
  anova_summary <- as.data.frame(do.call(rbind, anova_summary), stringsAsFactors = FALSE)
  anova_summary <- transform(anova_summary, p_value = as.numeric(p_value), f_value = as.numeric(f_value))
  return(anova_summary)
}

# To transpose
transpose_df <- function(df) {
  t_df <- data.table::transpose(df)
  colnames(t_df) <- rownames(df)
  rownames(t_df) <- colnames(df)
  t_df <- t_df %>%
    tibble::rownames_to_column(.data = .) %>%
    tibble::as_tibble(.)
  return(t_df)
}
```


### Loading Data

TCGA mRNA Expression

```{r}
brca_genes <- read_csv("../R/TCGA BRCA/mrna_top1000.csv", n_max = 1, col_names = FALSE)
brca_genes <- as.character(brca_genes[,2:1001])
brca_genes <- unlist(str_split(brca_genes, "\\|"))[ c(FALSE, TRUE) ]
```

METABRIC mRNA Expression
```{r}
metabric <- read_csv("../R/METABRIC/metabric_mrna_common_genes.csv")
metabric_labels <- read_csv("../R/METABRIC/PAM50_metabric.csv")
```
This data has been pre-processed to have only those features that were selected based on TCGA BRCA data beforehand. For details on the processing, check the file at /R/METABRIC/dataprocessing.Rmd.


### ANOVA

```{r}
anova_summary <- run_anova(metabric, metabric_labels, num_cpus=40)

median_pval <- median(anova_summary$p_value)
median_fval <- median(anova_summary$f_value)
```

## Random analysis

Here, I will select a random set of 1000 genes in the METABRIC mRNA Expression data and then find it's median F-value and p-value.

### Loading raw data
```{r}
colnames <- read_delim("../../metabric/EGAF00000102986/discovery_ExpressionMatrix.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE, n_max = 1, col_names = FALSE)
colnames <- c("genes", as.character(colnames[1,]))

discovery <- read_delim("../../metabric/EGAF00000102986/discovery_ExpressionMatrix.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE, col_names = colnames, skip = 1)

colnames <- read_delim("../../metabric/EGAF00000102987/validation_ExpressionMatrix.txt", 
    " ", escape_double = FALSE, trim_ws = TRUE, n_max = 1, col_names = FALSE)
colnames <- c("genes", as.character(colnames[1,]))

validation <- read_delim("../../metabric/EGAF00000102987/validation_ExpressionMatrix.txt", 
    " ", escape_double = FALSE, trim_ws = TRUE, col_names = colnames, skip = 1)
```

### Processing the data

The raw data needs quite a bit of processing to bring it to a more presentable format

Converting the Illumina Gene IDs to Gene Symbols
```{r}
# BiocManager::install("illuminaHumanv3.db")
library(illuminaHumanv3.db)
genes <- discovery$genes
genes <- data.frame(Gene=unlist(mget(x = genes, envir = illuminaHumanv3ENTREZREANNOTATED)))

discovery$genes <- genes$Gene

genes <- validation$genes
genes <- data.frame(Gene=unlist(mget(x = genes, envir = illuminaHumanv3ENTREZREANNOTATED)))

validation$genes <- genes$Gene
rm(genes)
```

Removing NAs and summarising over duplicates
```{r}
discovery <- discovery[complete.cases(discovery), ] %>%
      group_by(genes) %>%
      summarize(across(everything(), list(mean))) #%>%
#      remove_rownames() %>%
#      column_to_rownames("genes")

validation <- validation[complete.cases(validation), ] %>%
      group_by(genes) %>%
      summarize(across(everything(), list(mean))) #%>%
#      remove_rownames() %>%
#      column_to_rownames("genes")

```

Selecting only those features common to both discovery and validation sets
```{r}
common_genes <- intersect(discovery$genes, validation$genes)

metabric_d <- discovery %>%
  filter(genes %in% common_genes) %>%
  remove_rownames() %>%
  column_to_rownames("genes")

metabric_v <- validation %>%
  filter(genes %in% common_genes) %>%
  remove_rownames() %>%
  column_to_rownames("genes")

setdiff(metabric_d$genes, metabric_v$genes)
```

Transposing and then combining the discovery and validation sets
```{r}
metabric_d <- transpose_df(metabric_d) %>%
  dplyr::rename(patient_id = "rowname")
metabric_d$patient_id <- unlist(str_split(metabric_d$patient_id, "\\_"))[c(TRUE, FALSE)]

metabric_v <- transpose_df(metabric_v) %>%
  dplyr::rename(patient_id = "rowname")
metabric_v$patient_id <- unlist(str_split(metabric_v$patient_id, "\\_"))[c(TRUE, FALSE)]

metabric_complete <- rbind(metabric_d, metabric_v)
metabric_complete <- metabric_complete[match(metabric_labels$patient_id, metabric_complete$patient_id),]
```

### Selecting Random Features

```{r}
feature_list <- names(metabric_complete)[2:ncol(metabric_complete)]
set.seed(42)
random_sample <- sample(feature_list, 953)
# Adding "patient_id" column to random_sample
random_sample <- c("patient_id", random_sample)
```

### ANOVA

```{r}
anova_summary_random <- metabric_complete %>%
  dplyr::select(random_sample) %>%
  run_anova(metabric_labels, num_cpus=20)

median_pval_random <- median(anova_summary_random$p_value)
median_fval_random <- median(anova_summary_random$f_value)
```

## Comparison

```{r}
summary(anova_summary$p_value)
summary(anova_summary_random$p_value)
median_pval
median_pval_random

summary(anova_summary$f_value)
summary(anova_summary_random$f_value)
median_fval
median_fval_random
```

```{r}
df <- tibble(
  type = c(rep("tcga", nrow(anova_summary)), rep("random", nrow(anova_summary_random))),
  f_value = c(anova_summary$f_value, anova_summary_random$f_value),
  p_value = c(anova_summary$p_value, anova_summary_random$p_value)
)
```

```{r}
ggplot(df, aes(x=p_value, fill=type))+
  geom_density(alpha=0.5)+
  scale_x_continuous(trans='log10')

ggplot(df, aes(x=f_value, fill=type))+
  geom_density(alpha=0.5)+
  scale_x_continuous(trans='log10')
```



```{r eval=FALSE, include=FALSE}
## Selecting best features from METABRIC
anova_summary_metabric <- run_anova(metabric_complete, metabric_labels, num_cpus=40)
anova_metabric_sorted <- anova_summary_metabric %>% 
                  filter(p_value<=0.0001/nrow(anova_summary_metabric)) %>% 
                  slice_max(order_by=f_value, n=953)

ggplot(anova_metabric_sorted, aes(x=p_value))+
  geom_density(alpha=0.5)+
  scale_x_continuous(trans='log10')
```

```{r eval=FALSE, include=FALSE}
df <- tibble(
  type = c(rep("tcga", nrow(anova_summary)), 
           rep("random", nrow(anova_summary_random)),
           rep("metabric", nrow(anova_metabric_sorted))),
  f_value = c(anova_summary$f_value, anova_summary_random$f_value, anova_metabric_sorted$f_value),
  p_value = c(anova_summary$p_value, anova_summary_random$p_value, anova_metabric_sorted$p_value)
)

df %>%
  filter(p_value != 0) %>%
  ggplot(aes(x=p_value, fill=type))+
  geom_density(alpha=0.5)+
  scale_x_log10()
  scale_x_continuous(trans='log10')

ggplot(df, aes(x=f_value, fill=type))+
  geom_density(alpha=0.5)+
  scale_x_continuous(trans='log10')
```

```{r eval=FALSE, include=FALSE}
metabric <- intersect(as.character(discovery$genes), as.character(validation$genes))

tcga <- read_delim("/data/users/bs16b001/firehose/stddata__2016_01_28/BRCA/20160128/gdac.broadinstitute.org_BRCA.mRNAseq_Preprocess.Level_3.2016012800.0.0/BRCA.uncv2.mRNAseq_RSEM_normalized_log2.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
tcga <- tcga$gene
tcga <- unlist(str_split(tcga, "\\|"))[ c(FALSE, TRUE) ]

length(intersect(metabric, tcga))

tcga_selected <- unlist(str_split(brca_genes, "\\|"))[ c(FALSE, TRUE) ]
metabric_selected <- metabric_best1000

length(setdiff(tcga_selected, intersect(metabric, tcga)))
length(setdiff(metabric_selected, intersect(metabric, tcga)))

# Is this overlap significant?
q = 94              # Size of the overlap
m = 1000 - 47       # Number of genes selected in TCGA - genes not present in METABRIC
k = 1000 - 17       # Number of genes selected in METABRIC - genes not present in TCGA
n = 17206           # Total number of common genes in TCGA and METABRIC

phyper(q-1, m, n-m, k, lower.tail = FALSE, log.p = FALSE)
```

