---
title: "Data Preliminaries"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggfortify)

# To transpose
transpose_df <- function(df) {
  t_df <- data.table::transpose(df)
  colnames(t_df) <- rownames(df)
  rownames(t_df) <- colnames(df)
  t_df <- t_df %>%
    tibble::rownames_to_column(.data = .) %>%
    tibble::as_tibble(.)
  return(t_df)
}
```


# Checking for Covariates

## Loading data
```{r message=FALSE, warning=FALSE}
mrna <- read_csv("mrna_top1000.csv")
meth <- read_csv("meth_top1000.csv")
mirna <- read_csv("mirna_anova.csv")
pam50 <- read_csv("PAM50_subtype.csv")

data <- left_join(left_join(mrna, meth, by = "patient_id"), mirna, by = "patient_id")
data <- left_join(data, pam50, "patient_id")

data$patient_id <- str_sub(data$patient_id, 1, 12)
```

## Loading metadata and making it presentable
```{r message=FALSE, warning=FALSE}
BRCA_clin_merged_picked <- read_delim("../../../firehose/stddata__2016_01_28/BRCA/20160128/gdac.broadinstitute.org_BRCA.Clinical_Pick_Tier1.Level_4.2016012800.0.0/BRCA.clin.merged.picked.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

clinical1 <- BRCA_clin_merged_picked %>% 
              column_to_rownames(var="Hybridization REF") %>%
              transpose_df() %>%
              dplyr::rename("patient_id" = "rowname")

All_CDEs <- read_delim("../../../firehose/stddata__2016_01_28/BRCA/20160128/gdac.broadinstitute.org_BRCA.Clinical_Pick_Tier1.Level_4.2016012800.0.0/All_CDEs.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

clinical2 <- All_CDEs %>% 
              column_to_rownames(var="bcr_patient_barcode") %>%
              transpose_df() %>%
              dplyr::select(-patient_id) %>%
              dplyr::rename("patient_id" = "rowname")

clinical <- inner_join(clinical1, clinical2, by = "patient_id") 
#dplyr::select(intersect(names(clinical1), names(clinical2)))

clinical$patient_id <- str_to_upper(clinical$patient_id)
```

## Obtaining TSS from  Patient Barcode
TSS = Tissue Source Site
```{r}
data$TSS <- str_sub(data$patient_id, 6, 7)
```



## Filtering data

```{r}
patients <- intersect(data$patient_id, clinical$patient_id)

# Data
data <- data %>%
  dplyr::filter(patient_id %in% patients)

# Clinical data
clinical <- clinical %>%
  dplyr:: filter(patient_id %in% patients)

meta_and_data <- left_join(data, clinical, by = "patient_id") %>%
  dplyr::rename("Subtype" = "cancer_subtype",
                "Age" = "age_at_initial_pathologic_diagnosis",
                "Race" = "race.x",
                "Ethnicity" = "ethnicity.x") %>%
  mutate(Age = cut(as.numeric(Age), c(0,40,50,60,70,80,100)),
         Age = fct_recode(Age, "<40" = "(0,40]", ">80" = "(80,100]"),
         TSS = fct_lump_n(TSS, n=9),
         Race = gsub("black or african american", "black", Race),
         Race = str_to_title(Race),
         Ethnicity = gsub("hispanic or latino", "hispanic", Ethnicity),
         Ethnicity = gsub("not hispanic or latino", "not hispanic", Ethnicity),
         Ethnicity = str_to_title(Ethnicity))
```


## Performing Principal Component Analysis

```{r}
tcga.pca <- prcomp(data[,c(2:2258)], center = TRUE, scale. = TRUE)
#summary(tcga.pca)
str(tcga.pca)
```

## Checking for covairates


### Tissue Source Site
```{r}
autoplot(tcga.pca, data=meta_and_data, colour="TSS", frame=TRUE, frame.type="norm") +
  labs(title = "PCA on scaled data, coloured based on Tissue Source Site")
```

### Sequencing Center

All tissue samples were taken from cold storage and then sequenced in the same center. Different omics data was generated at different centers, but all samples' individual omics was generated at the same center.  

### Cancer Subtype
```{r}
autoplot(tcga.pca, data=meta_and_data, colour="Subtype", frame=TRUE, frame.type="norm") +
  labs(title = "PCA on scaled data, coloured based on Breast Cancer Subtype")
```

### Age
```{r}
autoplot(tcga.pca, data=meta_and_data, colour="Age", frame=TRUE, frame.type="norm") +
  labs(colour = "Age (in yrs)", fill = "Age (in yrs)") +
  labs(title = "PCA on scaled data, coloured based on Age")
```

### Race
```{r}
autoplot(tcga.pca, data=meta_and_data, colour="Race", frame=TRUE, frame.type="norm") +
  labs(title = "PCA on scaled data, coloured based on Race")
```

### Ethnicity
```{r}
autoplot(tcga.pca, data=meta_and_data, colour="Ethnicity", frame=TRUE, frame.type="norm") +
  labs(title = "PCA on scaled data, coloured based on Ethnicity")
```


# Scaling and Centering data

### mRNA Expression
```{r message=FALSE, warning=FALSE}
data <- read_csv("mrna_top1000.csv")
trte <- read.csv("trte_partition.txt", header=FALSE, colClasses = "character")
train <- as.character(trte[2,])
train <- train[train != ""]
test <- as.character(trte[4,])
test <- test[test != ""]

data_train <- filter(data, patient_id %in% train)
data_test <- filter(data, patient_id %in% test)

mean <- apply(data_train[2:1001], 2, mean)
std <- apply(data_train[2:1001], 2, sd)

data_train[2:1001] <- scale(data_train[2:1001], center = mean, scale = std)
data_test[2:1001] <- scale(data_test[2:1001], center = mean, scale = std)

new_data <- rbind(data_train, data_test)
data <- new_data[match(data$patient_id, new_data$patient_id),]
write_csv(data, "mrna_top1000_scaled.csv")
```

### DNA Methylation
```{r message=FALSE, warning=FALSE}
data <- read_csv("meth_top1000.csv")
trte <- read.csv("trte_partition.txt", header=FALSE, colClasses = "character")
train <- as.character(trte[2,])
train <- train[train != ""]
test <- as.character(trte[4,])
test <- test[test != ""]

data_train <- filter(data, patient_id %in% train)
data_test <- filter(data, patient_id %in% test)

mean <- apply(data_train[2:1001], 2, mean)
std <- apply(data_train[2:1001], 2, sd)

data_train[2:1001] <- scale(data_train[2:1001], center = mean, scale = std)
data_test[2:1001] <- scale(data_test[2:1001], center = mean, scale = std)

new_data <- rbind(data_train, data_test)
data <- new_data[match(data$patient_id, new_data$patient_id),]
write_csv(data, "meth_top1000_scaled.csv")
```

### miRNA Expression
```{r message=FALSE, warning=FALSE}
data <- read_csv("mirna_anova.csv")
trte <- read.csv("trte_partition.txt", header=FALSE, colClasses = "character")
train <- as.character(trte[2,])
train <- train[train != ""]
test <- as.character(trte[4,])
test <- test[test != ""]

data_train <- filter(data, patient_id %in% train)
data_test <- filter(data, patient_id %in% test)

mean <- apply(data_train[2:258], 2, mean)
std <- apply(data_train[2:258], 2, sd)

data_train[2:258] <- scale(data_train[2:258], center = mean, scale = std)
data_test[2:258] <- scale(data_test[2:258], center = mean, scale = std)

new_data <- rbind(data_train, data_test)
data <- new_data[match(data$patient_id, new_data$patient_id),]
write_csv(data, "mirna_anova_scaled.csv")
```

### mRNA Expression X mRNA Expression
```{r message=FALSE, warning=FALSE}
data <- read_csv("mrna_mrna_anova.csv")
trte <- read.csv("trte_partition.txt", header=FALSE, colClasses = "character")
train <- as.character(trte[2,])
train <- train[train != ""]
test <- as.character(trte[4,])
test <- test[test != ""]

data_train <- filter(data, patient_id %in% train)
data_test <- filter(data, patient_id %in% test)

mean <- apply(data_train[2:1001], 2, mean)
std <- apply(data_train[2:1001], 2, sd)

data_train[2:1001] <- scale(data_train[2:1001], center = mean, scale = std)
data_test[2:1001] <- scale(data_test[2:1001], center = mean, scale = std)

new_data <- rbind(data_train, data_test)
data <- new_data[match(data$patient_id, new_data$patient_id),]
write_csv(data, "mrna_mrna_anova_scaled.csv")
```

### DNA Methylation X DNA Methylation
```{r message=FALSE, warning=FALSE}
data <- read_csv("meth_meth_anova.csv")
trte <- read.csv("trte_partition.txt", header=FALSE, colClasses = "character")
train <- as.character(trte[2,])
train <- train[train != ""]
test <- as.character(trte[4,])
test <- test[test != ""]

data_train <- filter(data, patient_id %in% train)
data_test <- filter(data, patient_id %in% test)

mean <- apply(data_train[2:1001], 2, mean)
std <- apply(data_train[2:1001], 2, sd)

data_train[2:1001] <- scale(data_train[2:1001], center = mean, scale = std)
data_test[2:1001] <- scale(data_test[2:1001], center = mean, scale = std)

new_data <- rbind(data_train, data_test)
data <- new_data[match(data$patient_id, new_data$patient_id),]
write_csv(data, "meth_meth_anova_scaled.csv")
```

### miRNA Expression X miRNA Expression
```{r message=FALSE, warning=FALSE}
data <- read_csv("mirna_mirna_anova.csv")
trte <- read.csv("trte_partition.txt", header=FALSE, colClasses = "character")
train <- as.character(trte[2,])
train <- train[train != ""]
test <- as.character(trte[4,])
test <- test[test != ""]

data_train <- filter(data, patient_id %in% train)
data_test <- filter(data, patient_id %in% test)

mean <- apply(data_train[2:630], 2, mean)
std <- apply(data_train[2:630], 2, sd)

data_train[2:630] <- scale(data_train[2:630], center = mean, scale = std)
data_test[2:630] <- scale(data_test[2:630], center = mean, scale = std)

new_data <- rbind(data_train, data_test)
data <- new_data[match(data$patient_id, new_data$patient_id),]
write_csv(data, "mirna_mirna_anova_scaled.csv")
```

### mRNA Expression X DNA Methylation
```{r message=FALSE, warning=FALSE}
data <- read_csv("meth_mrna_anova.csv")
trte <- read.csv("trte_partition.txt", header=FALSE, colClasses = "character")
train <- as.character(trte[2,])
train <- train[train != ""]
test <- as.character(trte[4,])
test <- test[test != ""]

data_train <- filter(data, patient_id %in% train)
data_test <- filter(data, patient_id %in% test)

mean <- apply(data_train[2:1001], 2, mean)
std <- apply(data_train[2:1001], 2, sd)

data_train[2:1001] <- scale(data_train[2:1001], center = mean, scale = std)
data_test[2:1001] <- scale(data_test[2:1001], center = mean, scale = std)

new_data <- rbind(data_train, data_test)
data <- new_data[match(data$patient_id, new_data$patient_id),]
write_csv(data, "meth_mrna_anova_scaled.csv")
```

### DNA Methylation X miRNA Expression
```{r message=FALSE, warning=FALSE}
data <- read_csv("meth_mirna_anova.csv")
trte <- read.csv("trte_partition.txt", header=FALSE, colClasses = "character")
train <- as.character(trte[2,])
train <- train[train != ""]
test <- as.character(trte[4,])
test <- test[test != ""]

data_train <- filter(data, patient_id %in% train)
data_test <- filter(data, patient_id %in% test)

mean <- apply(data_train[2:1001], 2, mean)
std <- apply(data_train[2:1001], 2, sd)

data_train[2:1001] <- scale(data_train[2:1001], center = mean, scale = std)
data_test[2:1001] <- scale(data_test[2:1001], center = mean, scale = std)

new_data <- rbind(data_train, data_test)
data <- new_data[match(data$patient_id, new_data$patient_id),]
write_csv(data, "meth_mirna_anova_scaled.csv")
```

### miRNA Expression X mRNA Expression
```{r message=FALSE, warning=FALSE}
data <- read_csv("mrna_mirna_anova.csv")
trte <- read.csv("trte_partition.txt", header=FALSE, colClasses = "character")
train <- as.character(trte[2,])
train <- train[train != ""]
test <- as.character(trte[4,])
test <- test[test != ""]

data_train <- filter(data, patient_id %in% train)
data_test <- filter(data, patient_id %in% test)

mean <- apply(data_train[2:802], 2, mean)
std <- apply(data_train[2:802], 2, sd)

data_train[2:802] <- scale(data_train[2:802], center = mean, scale = std)
data_test[2:802] <- scale(data_test[2:802], center = mean, scale = std)

new_data <- rbind(data_train, data_test)
data <- new_data[match(data$patient_id, new_data$patient_id),]
write_csv(data, "mrna_mirna_anova_scaled.csv")
```