---
title: "Splitting Data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Importing the necessary the packages:
```{r}
library(tidyverse)
library(data.table)
library(readr)
library(broom)
library(caret)
```

## Importing data
First we import just the samples from each data type to find the common samples, and then we will use use them and their labels to do a stratified random train test split.

### DNA Methylation
```{r}
meth <- read_delim("/data/users/bs16b001/firehose/stddata__2016_01_28/BRCA/20160128/gdac.broadinstitute.org_BRCA.Merge_methylation__humanmethylation450__jhu_usc_edu__Level_3__within_bioassay_data_set_function__data.Level_3.2016012800.0.0/meth.txt", 
      "\t", escape_double = FALSE, trim_ws = TRUE, n_max=1, col_names = FALSE)
meth <- unique(unlist(meth)[2:length(meth)])

#Below is the format for the sample names. This removes the aliquot id etc
pattern <- "^(\\w{4}\\-\\w{2}\\-\\w{4}\\-\\w{2}).*"

#Now to get meth samples in that format
meth <- sub(pattern, "\\1", meth)
```

### mRNA Expression
```{r}
mrna <- read_delim("/data/users/bs16b001/firehose/stddata__2016_01_28/BRCA/20160128/gdac.broadinstitute.org_BRCA.mRNAseq_Preprocess.Level_3.2016012800.0.0/BRCA.uncv2.mRNAseq_RSEM_normalized_log2.txt", 
    "\t", n_max=1, col_names = FALSE)
mrna <- unique(unlist(mrna)[2:length(mrna)])

#mrna sample ids already follow the format. If not, uncomment the next line and run it.
#mrna <- sub(pattern, "\\1", mrna)
```

### microRNA Expression
```{r}
mirna <- read_delim("/data/users/bs16b001/firehose/stddata__2016_01_28/BRCA/20160128/gdac.broadinstitute.org_BRCA.Merge_mirnaseq__illuminahiseq_mirnaseq__bcgsc_ca__Level_3__miR_isoform_expression__data.Level_3.2016012800.0.0/mirna.txt", 
     "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
          dplyr::select(SampleId)
mirna <-  unique(mirna$SampleId)

#Now to get mirna samples in the right format
mirna <- sub(pattern, "\\1", mirna)
```

### Determining the common samples
```{r}
common_samples <- intersect(intersect(meth, mrna), mirna)
```

## Adding labels and splitting the data

### Reading and adding labels
First, we have to read the file containing the PAM50 labels
```{r}
pam50 <- read_csv("/data/users/bs16b001/R/pam50.csv") %>%
          mutate(Sample_ID = sub(pattern, "\\1", Sample_ID)) %>%
          filter(Sample_ID %in% common_samples)

#Now since for some samples the PAM50 label was not present in the file, we will remove them from the common samples list too.
common_samples <- unlist(pam50["Sample_ID"], use.names = FALSE)
```

### Stratified Random Split
```{r}
train.index <- createDataPartition(pam50$PAM50, p = .7, list = FALSE)
train <- pam50[ train.index,]
test  <- pam50[-train.index,]
```

## Writing files
### The PAM50_subtype file
```{r}
pam50 %>%
  dplyr::rename(patient_id = "Sample_ID", cancer_subtype = "PAM50")# %>%
  #write.csv("PAM50_subtype.csv", row.names = FALSE)
```

## The train-test partition file
```{r}
#fileConn<-file("trte_partition.txt")
#writeLines(c("training patient_id", paste(unlist(train["Sample_ID"]), collapse = ","), "testing patient_id", paste(unlist(test["Sample_ID"]), collapse = ",")), fileConn)
#close(fileConn)
```

## The distribution
```{r}
pam50$Category <- ifelse(pam50$Sample_ID %in% train$Sample_ID, "Train", "Test")

pam50 %>%
  ggplot(aes(x=PAM50, group=Category, fill=Category))+
  geom_bar(position="dodge")+
  labs(x="Cancer subtype", y="Number of samples", title = "Sample distribution")

(Number_of_train_samples = length(train.index))
(Number_of_test_samples = nrow(pam50) - length(train.index))
(Total_samples = nrow(pam50))
```

