---
title: "Interactions"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Importing all the requisite packages first.
```{r message=FALSE}
library(tidyverse)
library(data.table)
library(readr)
library(broom)
library(caret)
library(foreach)
library(doParallel)
```
```{r custom functions}
custom_anova <- function(data, labels) {
  num_cpus <- 40
  cl <- makeCluster(num_cpus)
  doParallel::registerDoParallel(cl)
  anova_summary <- foreach(i = 3:ncol(data)) %dopar% {
    column <- names(data[i]) 
    avz <- broom:: tidy(aov(data[,i] ~ labels)) # Each ANOVA test iterating through each column
    return(c(gene=column, f_value=avz$statistic[[1]], p_value=avz$p.value[[1]]))
  }
  stopCluster(cl)
  anova_summary <- as.data.frame(do.call(rbind, anova_summary), stringsAsFactors = FALSE)
  anova_summary <- transform(anova_summary, p_value = as.numeric(p_value), f_value = as.numeric(f_value))

  #Filter the features according to the ANOVA p-value and then sort them according to the f_values, which is not necessary here, but can be useful for functional and representational purposes later
  sorted_data <- anova_summary %>% 
                  filter(p_value<=0.001/nrow(anova_summary)) %>% 
                  arrange(desc(f_value))
  return(sorted_data)
}

greedySelect <- function(dframe, featureList, corrThreshold = 0.8, max_features = 1000){
  selected = c(featureList[1])
    for (i in c(2:length(featureList))) {
      tempList <- c(cor(dframe[featureList[i]], dframe[selected[1]])) 
      for (s in c(1:length(selected))){
        temp <- cor(dframe[featureList[i]], dframe[selected[s]])
        tempList <- append(tempList, temp)
        if (abs(temp) > corrThreshold) {break}
        }
      if (all( abs(tempList) < corrThreshold)) {
        selected <- append(selected, featureList[i])
      }
      if (length(selected) == max_features) {break}
    }
  return(selected)
  }
```

# Self-interactions

## miRNA expression

Reading in the data.
```{r}
mirna <- read_csv("mirna_complete.csv")
pam50 <- read_csv("PAM50_subtype.csv")

# Getting the training indices
trte <- file("trte_partition.txt", open = "r")
lines <- readLines(trte)
close(trte)
train_idx <- unlist(strsplit(lines[2], ","))

# Selecting only training samples for doing the processing
mirna_train <- filter(mirna, mirna$patient_id %in% train_idx)
pam50_train <- filter(pam50, pam50$patient_id %in% train_idx)

#Check that the labels are identical
identical(pam50_train$patient_id, mirna_train$patient_id) #This has to be true
```

Finding the self-interactions.
```{r}
mirna_train <- column_to_rownames(mirna_train, var = "patient_id")
col_names <- sapply(combn(colnames(mirna_train), 2, simplify = FALSE), paste, collapse = "__")
mirna_mirna_train <- setNames(data.frame(combn(mirna_train, 2, FUN = Reduce, f = `*`)), col_names)
row.names(mirna_mirna_train) <- row.names(mirna_train)
```
## ANOVA
```{r}
num_clusters <- 40
cl <- makeCluster(num_clusters)
doParallel::registerDoParallel(cl)
anova_summary <- foreach(i = 3:ncol(mirna_mirna_train)) %dopar% {
  column <- names(mirna_mirna_train[i]) # to print each ROI at the top of each ANOVA summary
  avz <- broom:: tidy(aov(mirna_mirna_train[,i] ~ unlist(pam50_train$cancer_subtype))) # Each ANOVA test iterating through each column
  return(c(microRNA=column, f_value=avz$statistic[[1]], p_value=avz$p.value[[1]]))
}
stopCluster(cl)
anova_summary <- as.data.frame(do.call(rbind, anova_summary), stringsAsFactors = FALSE)
anova_summary <- transform(anova_summary, p_value = as.numeric(p_value), f_value = as.numeric(f_value))
```

Filter the features according to the ANOVA p-value and then sort them according to the f_values, which is not necessary here, but can be useful for functional and representational purposes later
```{r}
mirna_mirna_sorted <- anova_summary %>% 
                  filter(p_value<(0.001/nrow(anova_summary))) %>% 
                  arrange(desc(f_value))
```
Now to try a better approach to get the best features: A greedy approach to selecting the top non-correlated features
```{r}
greedySelect <- function(dframe, featureList, corrThreshold = 0.8, max_features = 1000){
  selected = c(featureList[1])
    for (i in c(2:length(featureList))) {
      tempList <- c(cor(dframe[featureList[i]], dframe[selected[1]])) 
      for (s in c(1:length(selected))){
        temp <- cor(dframe[featureList[i]], dframe[selected[s]])
        tempList <- append(tempList, temp)
        if (abs(temp) > corrThreshold) {break}
        }
      if (all( abs(tempList) < corrThreshold)) {
        selected <- append(selected, featureList[i])
      }
      if (length(selected) == max_features) {break}
    }
  return(selected)
  }
```

```{r}
mirna_mirna_greedy_names <- greedySelect(mirna_mirna_train, mirna_mirna_sorted["microRNA"][[1]], corrThreshold = 0.8)

mirna <- column_to_rownames(mirna, var = "patient_id")
col_names <- sapply(combn(colnames(mirna), 2, simplify = FALSE), paste, collapse = "__")
mirna_mirna <- setNames(data.frame(combn(mirna, 2, FUN = Reduce, f = `*`)), col_names)
row.names(mirna_mirna) <- row.names(mirna)

index <- match(mirna_mirna_greedy_names, names(mirna_mirna))
mirna_mirna_greedy_df <- mirna_mirna[,index]
row.names(mirna_mirna_greedy_df) <- row.names(mirna)
mirna_mirna_greedy_df <- mirna_mirna_greedy_df %>%
                        rownames_to_column(var = "patient_id")
write.csv(mirna_mirna_greedy_df, "mirna_mirna_anova.csv", row.names = FALSE)
```

## mRNA expression

Reading in the data.
```{r}
mrna <- read_csv("mrna_top1000.csv")
pam50 <- read_csv("PAM50_subtype.csv")

# Getting the training indices
trte <- file("trte_partition.txt", open = "r")
lines <- readLines(trte)
close(trte)
train_idx <- unlist(strsplit(lines[2], ","))

# Selecting only training samples for doing the processing
mrna_train <- filter(mrna, mrna$patient_id %in% train_idx)
pam50_train <- filter(pam50, pam50$patient_id %in% train_idx)

#Check that the labels are identical
identical(pam50_train$patient_id, mrna_train$patient_id) #This has to be true
```

Finding the self-interactions.
```{r}
mrna_train <- column_to_rownames(mrna_train, var = "patient_id")
col_names <- sapply(combn(colnames(mrna_train), 2, simplify = FALSE), paste, collapse = "__")
mrna_mrna_train <- setNames(data.frame(combn(mrna_train, 2, FUN = Reduce, f = `*`)), col_names)
row.names(mrna_mrna_train) <- row.names(mrna_train)
```
## ANOVA
```{r}
num_clusters <- 40
cl <- makeCluster(num_clusters)
doParallel::registerDoParallel(cl)
anova_summary <- foreach(i = 3:ncol(mrna_mrna_train)) %dopar% {
  column <- names(mrna_mrna_train[i]) # to print each ROI at the top of each ANOVA summary
  avz <- broom:: tidy(aov(mrna_mrna_train[,i] ~ unlist(pam50_train$cancer_subtype))) # Each ANOVA test iterating through each column
  return(c(gene=column, f_value=avz$statistic[[1]], p_value=avz$p.value[[1]]))
}
stopCluster(cl)
anova_summary <- as.data.frame(do.call(rbind, anova_summary), stringsAsFactors = FALSE)
anova_summary <- transform(anova_summary, p_value = as.numeric(p_value), f_value = as.numeric(f_value))
```

Filter the features according to the ANOVA p-value and then sort them according to the f_values, which is not necessary here, but can be useful for functional and representational purposes later
```{r}
mrna_mrna_sorted <- anova_summary %>% 
                  filter(p_value<(0.001/nrow(anova_summary))) %>% 
                  arrange(desc(f_value))
```
Now to try a better approach to get the best features: A greedy approach to selecting the top non-correlated features
```{r}
greedySelect <- function(dframe, featureList, corrThreshold = 0.8, max_features = 1000){
  selected = c(featureList[1])
    for (i in c(2:length(featureList))) {
      tempList <- c(cor(dframe[featureList[i]], dframe[selected[1]])) 
      for (s in c(1:length(selected))){
        temp <- cor(dframe[featureList[i]], dframe[selected[s]])
        tempList <- append(tempList, temp)
        if (abs(temp) > corrThreshold) {break}
        }
      if (all( abs(tempList) < corrThreshold)) {
        selected <- append(selected, featureList[i])
      }
      if (length(selected) == max_features) {break}
    }
  return(selected)
  }
```

```{r}
mrna_mrna_greedy_names <- greedySelect(mrna_mrna_train, mrna_mrna_sorted["gene"][[1]], corrThreshold = 0.8)

mrna <- column_to_rownames(mrna, var = "patient_id")
col_names <- sapply(combn(colnames(mrna), 2, simplify = FALSE), paste, collapse = "__")
mrna_mrna <- setNames(data.frame(combn(mrna, 2, FUN = Reduce, f = `*`)), col_names)
row.names(mrna_mrna) <- row.names(mrna)

index <- match(mrna_mrna_greedy_names, names(mrna_mrna))
mrna_mrna_greedy_df <- mrna_mrna[,index]
row.names(mrna_mrna_greedy_df) <- row.names(mrna)
mrna_mrna_greedy_df <- mrna_mrna_greedy_df %>%
                        rownames_to_column(var = "patient_id")
write.csv(mrna_mrna_greedy_df, "mrna_mrna_anova.csv", row.names = FALSE)

rm(list = ls())
```


## DNA Methylation

Reading in the data.
```{r}
meth <- read_csv("meth_top1000.csv")
pam50 <- read_csv("PAM50_subtype.csv")

# Getting the training indices
trte <- file("trte_partition.txt", open = "r")
lines <- readLines(trte)
close(trte)
train_idx <- unlist(strsplit(lines[2], ","))

# Selecting only training samples for doing the processing
meth_train <- filter(meth, meth$patient_id %in% train_idx)
pam50_train <- filter(pam50, pam50$patient_id %in% train_idx)

#Check that the labels are identical
identical(pam50_train$patient_id, meth_train$patient_id) #This has to be true
```

Finding the self-interactions.
```{r}
meth_train <- column_to_rownames(meth_train, var = "patient_id")
col_names <- sapply(combn(colnames(meth_train), 2, simplify = FALSE), paste, collapse = "__")
meth_meth_train <- setNames(data.frame(combn(meth_train, 2, FUN = Reduce, f = `*`)), col_names)
row.names(meth_meth_train) <- row.names(meth_train)
```
## ANOVA
```{r}
num_clusters <- 40
cl <- makeCluster(num_clusters)
doParallel::registerDoParallel(cl)
anova_summary <- foreach(i = 3:ncol(meth_meth_train)) %dopar% {
  column <- names(meth_meth_train[i]) # to print each ROI at the top of each ANOVA summary
  avz <- broom:: tidy(aov(meth_meth_train[,i] ~ unlist(pam50_train$cancer_subtype))) # Each ANOVA test iterating through each column
  return(c(gene=column, f_value=avz$statistic[[1]], p_value=avz$p.value[[1]]))
}
stopCluster(cl)
anova_summary <- as.data.frame(do.call(rbind, anova_summary), stringsAsFactors = FALSE)
anova_summary <- transform(anova_summary, p_value = as.numeric(p_value), f_value = as.numeric(f_value))
```

Filter the features according to the ANOVA p-value and then sort them according to the f_values, which is not necessary here, but can be useful for functional and representational purposes later
```{r}
meth_meth_sorted <- anova_summary %>% 
                  filter(p_value<(0.001/nrow(anova_summary))) %>% 
                  arrange(desc(f_value))
```
Now to try a better approach to get the best features: A greedy approach to selecting the top non-correlated features
```{r}
greedySelect <- function(dframe, featureList, corrThreshold = 0.8, max_features = 1000){
  selected = c(featureList[1])
    for (i in c(2:length(featureList))) {
      tempList <- c(cor(dframe[featureList[i]], dframe[selected[1]])) 
      for (s in c(1:length(selected))){
        temp <- cor(dframe[featureList[i]], dframe[selected[s]])
        tempList <- append(tempList, temp)
        if (abs(temp) > corrThreshold) {break}
        }
      if (all( abs(tempList) < corrThreshold)) {
        selected <- append(selected, featureList[i])
      }
      if (length(selected) == max_features) {break}
    }
  return(selected)
  }
```

```{r}
meth_meth_greedy_names <- greedySelect(meth_meth_train, meth_meth_sorted["gene"][[1]], corrThreshold = 0.8)

meth <- column_to_rownames(meth, var = "patient_id")
col_names <- sapply(combn(colnames(meth), 2, simplify = FALSE), paste, collapse = "__")
meth_meth <- setNames(data.frame(combn(meth, 2, FUN = Reduce, f = `*`)), col_names)
row.names(meth_meth) <- row.names(meth)

index <- match(meth_meth_greedy_names, names(meth_meth))
meth_meth_greedy_df <- meth_meth[,index]
row.names(meth_meth_greedy_df) <- row.names(meth)
meth_meth_greedy_df <- meth_meth_greedy_df %>%
                        rownames_to_column(var = "patient_id")
write.csv(meth_meth_greedy_df, "meth_meth_anova.csv", row.names = FALSE)
```

# Cross-interactions

Reading in all the data.
```{r}
pam50 <- read_csv("PAM50_subtype.csv")
# Getting the training indices
trte <- file("trte_partition.txt", open = "r")
lines <- readLines(trte)
close(trte)
train_idx <- unlist(strsplit(lines[2], ","))

pam50_train <- filter(pam50, pam50$patient_id %in% train_idx)


meth <- read_csv("meth_top1000.csv")
mrna <- read_csv("mrna_top1000.csv")
mirna <- read_csv("mirna_complete.csv")


# Selecting only training samples for doing the processing
meth_train <- filter(meth, meth$patient_id %in% train_idx)
mrna_train <- filter(mrna, mrna$patient_id %in% train_idx)
mirna_train <- filter(mirna, mirna$patient_id %in% train_idx)

# Check that the labels are identical
identical(pam50_train$patient_id, meth_train$patient_id) #This has to be true
identical(pam50_train$patient_id, mrna_train$patient_id) #This has to be true
identical(pam50_train$patient_id, mirna_train$patient_id) #This has to be true

# Setting row.names
meth_train <- column_to_rownames(meth_train, var = "patient_id")
mrna_train <- column_to_rownames(mrna_train, var = "patient_id")
mirna_train <- column_to_rownames(mirna_train, var = "patient_id")
meth <- column_to_rownames(meth, var = "patient_id")
mrna <- column_to_rownames(mrna, var = "patient_id")
mirna <- column_to_rownames(mirna, var = "patient_id")
```

## DNA Methylation X mRNA Expression
Finding the interactions.
```{r warning=FALSE}
col_names <- apply(expand.grid(names(meth_train), names(mrna_train)), 1, paste, collapse="_")
meth_mrna_train <- setNames(as.data.frame(lapply(mrna_train, `*`, meth_train)), col_names)
row.names(meth_mrna_train) <- row.names(meth_train)

#ANOVA
meth_mrna_sorted <- custom_anova(meth_mrna_train, pam50_train$cancer_subtype)
```
```{r}
#Only selecting the top 10000 to be supplied
top25000 <- meth_mrna_train[, which(names(meth_mrna_train) %in% meth_mrna_sorted$gene[1:25000])]
meth_mrna_greedy_names <- greedySelect(top25000, meth_mrna_sorted["gene"][[1]][1:25000], corrThreshold = 0.9)

col_names <- apply(expand.grid(names(meth), names(mrna)), 1, paste, collapse="_")
meth_mrna <- setNames(as.data.frame(lapply(mrna, `*`, meth)), col_names)
row.names(meth_mrna) <- row.names(meth)

index <- match(meth_mrna_greedy_names, names(meth_mrna))
meth_mrna_greedy_df <- meth_mrna[,index]
row.names(meth_mrna_greedy_df) <- row.names(mrna)
meth_mrna_greedy_df <- meth_mrna_greedy_df %>%
                        rownames_to_column(var = "patient_id")
write.csv(meth_mrna_greedy_df, "meth_mrna_anova.csv", row.names = FALSE)
rm(meth_mrna, meth_mrna_train, meth_mrna_greedy_df)
```

## DNA Methylation X miRNA Expression
Finding the interactions.
```{r warning=FALSE}
col_names <- apply(expand.grid(names(meth_train), names(mirna_train)), 1, paste, collapse="_")
meth_mirna_train <- setNames(as.data.frame(lapply(mirna_train, `*`, meth_train)), col_names)
row.names(meth_mirna_train) <- row.names(meth_train)

#ANOVA
meth_mirna_sorted <- custom_anova(meth_mirna_train, pam50_train$cancer_subtype)

top25000 <- meth_mirna_train[, which(names(meth_mirna_train) %in% meth_mirna_sorted$gene[1:25000])]
meth_mirna_greedy_names <- greedySelect(top25000, meth_mirna_sorted["gene"][[1]][1:25000], corrThreshold = 0.95)

col_names <- apply(expand.grid(names(meth), names(mirna)), 1, paste, collapse="_")
meth_mirna <- setNames(as.data.frame(lapply(mirna, `*`, meth)), col_names)
row.names(meth_mirna) <- row.names(meth)

index <- match(meth_mirna_greedy_names, names(meth_mirna))
meth_mirna_greedy_df <- meth_mirna[,index]
row.names(meth_mirna_greedy_df) <- row.names(mirna)
meth_mirna_greedy_df <- meth_mirna_greedy_df %>%
                        rownames_to_column(var = "patient_id")
write.csv(meth_mirna_greedy_df, "meth_mirna_anova.csv", row.names = FALSE)
#rm(meth_mirna, meth_mirna_train, meth_mirna_greedy_df)
```

## mRNA Expression X miRNA Expression
Finding the interactions.
```{r warning=FALSE}
col_names <- apply(expand.grid(names(mrna_train), names(mirna_train)), 1, paste, collapse="_")
mrna_mirna_train <- setNames(as.data.frame(lapply(mirna_train, `*`, mrna_train)), col_names)
row.names(mrna_mirna_train) <- row.names(mrna_train)

#ANOVA
mrna_mirna_sorted <- custom_anova(mrna_mirna_train, pam50_train$cancer_subtype)
```
```{r}
top25000 <- mrna_mirna_train[, which(names(mrna_mirna_train) %in% mrna_mirna_sorted$gene[1:25000])]
mrna_mirna_greedy_names <- greedySelect(top25000, mrna_mirna_sorted["gene"][[1]][1:25000], corrThreshold = 0.95)

col_names <- apply(expand.grid(names(mrna), names(mirna)), 1, paste, collapse="_")
mrna_mirna <- setNames(as.data.frame(lapply(mirna, `*`, mrna)), col_names)
row.names(mrna_mirna) <- row.names(mrna)

index <- match(mrna_mirna_greedy_names, names(mrna_mirna))
mrna_mirna_greedy_df <- mrna_mirna[,index]
row.names(mrna_mirna_greedy_df) <- row.names(mirna)
mrna_mirna_greedy_df <- mrna_mirna_greedy_df %>%
                        rownames_to_column(var = "patient_id")
write.csv(mrna_mirna_greedy_df, "mrna_mirna_anova.csv", row.names = FALSE)
rm(mrna_mirna, mrna_mirna_train, mrna_mirna_greedy_df)
```
