---
title: "Simpler Methods"
output: html_notebook
---

Importing all the requisite packages first.
```{r message=FALSE}
library(tidyverse)
library(data.table)
library(readr)
library(broom)
library(caret)
library(foreach)
library(doParallel)
```

Custom functions
```{r}
f1_score <- function(predicted, true, type="weighted"){
  predicted <- factor(as.character(predicted), levels=unique(as.character(true)))
  true  <- factor(true, levels=unique(as.character(true)))
  cm = as.matrix(table(true, predicted))

  precision <- diag(cm) / colSums(cm)
  recall <- diag(cm) / rowSums(cm)
  f1 <-  ifelse(precision + recall == 0, 0, 2 * precision * recall / (precision + recall))

  #Assuming that F1 is zero when it's not possible compute it
  f1[is.na(f1)] <- 0

  #weighted average of the f1 scores
  weights <- table(true)/length(true)
  sum(f1*weights)
}

# pred <- c(0,0,0,0,0,0,0,0,0,0,0,0)
# true <- c(0,0,0,0,1,1,1,1,2,2,2,2)
# f1 <- f1_score(pred, true)
```


Reading in all the data.
```{r message=FALSE, warning=FALSE}
pam50 <- read_csv("PAM50_subtype.csv")
# Getting the training indices
trte <- file("trte_partition.txt", open = "r")
lines <- readLines(trte)
close(trte)
train_idx <- unlist(strsplit(lines[2], ","))
test_idx <- unlist(strsplit(lines[4], ","))

pam50_train <- filter(pam50, pam50$patient_id %in% train_idx)
pam50_test <- filter(pam50, pam50$patient_id %in% test_idx)


meth <- read_csv("meth_top1000.csv")
mrna <- read_csv("mrna_top1000.csv")
mirna <- read_csv("mirna_complete.csv")


# Selecting only training samples for model training
meth_train <- filter(meth, meth$patient_id %in% train_idx)
mrna_train <- filter(mrna, mrna$patient_id %in% train_idx)
mirna_train <- filter(mirna, mirna$patient_id %in% train_idx)

# Selecting only test samples for testing
meth_test <- filter(meth, meth$patient_id %in% test_idx)
mrna_test <- filter(mrna, mrna$patient_id %in% test_idx)
mirna_test <- filter(mirna, mirna$patient_id %in% test_idx)

# Check that the labels are identical
identical(pam50_train$patient_id, meth_train$patient_id) #This has to be true
identical(pam50_train$patient_id, mrna_train$patient_id) #This has to be true
identical(pam50_train$patient_id, mirna_train$patient_id) #This has to be true

# Setting row.names
meth_train <- column_to_rownames(meth_train, var = "patient_id")
mrna_train <- column_to_rownames(mrna_train, var = "patient_id")
mirna_train <- column_to_rownames(mirna_train, var = "patient_id")
meth_test <- column_to_rownames(meth_test, var = "patient_id")
mrna_test <- column_to_rownames(mrna_test, var = "patient_id")
mirna_test <- column_to_rownames(mirna_test, var = "patient_id")
meth <- column_to_rownames(meth, var = "patient_id")
mrna <- column_to_rownames(mrna, var = "patient_id")
mirna <- column_to_rownames(mirna, var = "patient_id")
```

Concatenating the data for easier use with most modelers.

```{r}
all_train <- bind_cols(mrna_train, meth_train, mirna_train)
all_test <- bind_cols(mrna_test, meth_test, mirna_test)
all_data <- bind_cols(mrna, meth, mirna)
```


# Modeling

## Linear modeling

Things to do:
- Note down these scores
- Put all this in notion
- Make the report "Why GCNs?"

### LASSO
Data formatting
```{r}
X <- scale(as.matrix(all_train[,1:ncol(all_train)]), center = TRUE, scale = TRUE)
Y <- as.factor(as.matrix(pam50_train$cancer_subtype))

X_test <- scale(as.matrix(all_test))
```

Modelling
```{r}
library(glmnet)
alpha <- 1

num_cpus <- 25
cl <- makeCluster(num_cpus) 
doParallel::registerDoParallel(cl)
cv <- foreach(alpha = alpha, .combine = rbind) %dopar% {
  glmnet::cv.glmnet(X, Y, family = "multinomial", nfold = 15, type.measure = "deviance", parallel = TRUE, alpha = alpha)
}
stopCluster(cl)

model <- glmnet(X, Y, family = "multinomial", lambda = cv$lambda.1se, alpha = alpha)
```
Evaluation
```{r}
train_predictions <- model %>% predict(X, type="class")
test_predictions <- model %>% predict(X_test, type="class")

train_f1 <- f1_score(train_predictions, pam50_train$cancer_subtype)
test_f1 <- f1_score(test_predictions, pam50_test$cancer_subtype)

print(train_f1)
print(test_f1)
# train_predictions %>%
#   as_tibble() %>%
#   group_by(s0) %>%
#   summarise(counts = n()) %>%
#   ggplot(aes(x=s0, y=counts))+
#   geom_bar(stat="identity")
```


### Ridge Regression
Data formatting
```{r}
X <- scale(as.matrix(all_train[,1:ncol(all_train)]), center = TRUE, scale = TRUE)
Y <- as.factor(as.matrix(pam50_train$cancer_subtype))

X_test <- scale(as.matrix(all_test))
```

Modelling
```{r}
library(glmnet)
alpha <- 0

cv <- glmnet::cv.glmnet(X, Y, family = "multinomial", nfold = 15, type.measure = "deviance", parallel = FALSE, alpha = alpha)

model <- glmnet(X, Y, family = "multinomial", lambda = cv$lambda.1se, alpha = alpha)
```

Evaluation
```{r}
train_predictions <- model %>% predict(X, type="class")
test_predictions <- model %>% predict(X_test, type="class")

train_f1 <- f1_score(train_predictions, pam50_train$cancer_subtype)
test_f1 <- f1_score(test_predictions, pam50_test$cancer_subtype)

print(train_f1)
print(test_f1)

# train_predictions %>%
#   as_tibble() %>%
#   group_by(s0) %>%
#   summarise(counts = n()) %>%
#   ggplot(aes(x=s0, y=counts))+
#   geom_bar(stat="identity")
```


### block_plsda and block_splsda
These are linear discriminant analyses with projection to latent structure models.
From the mixomics package.

Data formatting
```{r}
X1 <- scale(as.matrix(mrna_train[,1:ncol(mrna_train)]), center = TRUE, scale = TRUE)
X2 <- scale(as.matrix(meth_train[,1:ncol(meth_train)]), center = TRUE, scale = TRUE)
X3 <- scale(as.matrix(mirna_train[,1:ncol(mirna_train)]), center = TRUE, scale = TRUE)
X <- list(mrna=X1, meth=X2, mirna=X3)
Y <- as.factor(as.matrix(pam50_train$cancer_subtype))

X_test <- list(mrna=scale(as.matrix(mrna_test)), meth=scale(as.matrix(meth_test)), mirna=scale(as.matrix(mirna_test)))
```

Modelling - PLSDA
```{r}
# Projection to Latent Structures models with  Discriminant Analysis (PSLDA)
# library(mixOmics)

design <- matrix(c(0,1,1,1,0,1,1,1,0), ncol = 3, nrow = 3)

model <- block.plsda(X,Y, design=design, ncomp=5)
```

Evaluation
```{r}
# plotIndiv(model, ind.names = FALSE, legend = TRUE)
# plotVar(model, legend = TRUE)

train_predictions <- (model %>% predict(X, type="class"))$WeightedVote$centroids.dist[,2]
test_predictions <- (model %>% predict(X_test, type="class"))$WeightedVote$centroids.dist[,2]

train_f1 <- f1_score(train_predictions, pam50_train$cancer_subtype)
test_f1 <- f1_score(test_predictions, pam50_test$cancer_subtype)

print(train_f1)
print(test_f1)
```

Modelling - SPLSDA
```{r}
# Sparse Projection to Latent Structures models with  Discriminant Analysis (SPLSDA)
# library(mixOmics)

design <- matrix(c(1,1,1,1,1,1,1,1,1), ncol = 3, nrow = 3)

model <- block.splsda(X,Y, design=design, ncomp=5)
```

Evaluation
```{r}
# plotIndiv(model, ind.names = FALSE, legend = TRUE)
# plotVar(model, legend = TRUE)

train_predictions <- (model %>% predict(X, type="class"))$WeightedVote$centroids.dist[,2]
test_predictions <- (model %>% predict(X_test, type="class"))$WeightedVote$centroids.dist[,2]

train_f1 <- f1_score(train_predictions, pam50_train$cancer_subtype)
test_f1 <- f1_score(test_predictions, pam50_test$cancer_subtype)

print(train_f1)
print(test_f1)
```


## Machine Learning

### SVM

Data formatting
```{r}
X <- scale(as.matrix(all_train[,1:ncol(all_train)]), center = TRUE, scale = TRUE)
Y <- as.factor(as.matrix(pam50_train$cancer_subtype))

X_test <- scale(as.matrix(all_test))
```

Modelling and evaluation
```{r}
# library(e1071)

svmfit = svm(X, Y, kernel = "linear", cost = 10, scale = FALSE)

train_predictions <- svmfit %>% predict(X, type="class")
test_predictions <- svmfit %>% predict(X_test, type="class")

train_f1 <- f1_score(train_predictions, pam50_train$cancer_subtype)
test_f1 <- f1_score(test_predictions, pam50_test$cancer_subtype)

print(train_f1)
print(test_f1)
```



### Random Forest

Data formatting
```{r}
X <- scale(as.matrix(all_train[,1:ncol(all_train)]), center = TRUE, scale = TRUE)
Y <- as.factor(as.matrix(pam50_train$cancer_subtype))
X_test <- scale(as.matrix(all_test))
```

Modelling and evaluation
```{r}
# library(randomForest)

RF_model = randomForest(X, Y, kernel = "radial", cost = 10, scale = FALSE)

train_predictions <- RF_model %>% predict(X, type="class")
test_predictions <- RF_model %>% predict(X_test, type="class")

train_f1 <- f1_score(train_predictions, pam50_train$cancer_subtype)
test_f1 <- f1_score(test_predictions, pam50_test$cancer_subtype)

print(train_f1)
print(test_f1)
```