---
title: "Data Split and Label Recovery"
output: github_document
editor_options: 
  chunk_output_type: inline
---

In this notebook, I will get the labels from the data and then format them as needed and also split the data into training and testing splits.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(readr)
library(broom)
library(caret)
```

## Loading data
```{r message=FALSE, warning=FALSE}
data <- read_delim("../../../../HBTRC_data/GN326_MeanDataAnnotated_rev081815.txt", "\t", escape_double = FALSE, trim_ws = TRUE, skip = 32)
```
```{r message=FALSE, warning=FALSE}
# Getting the Gene Symbol and description and storing them as an annotation file
annotation <- data %>%
  select("Gene Symbol", "Gene Id", Description) %>%
  rename(Gene_Symbol = "Gene Symbol", GeneID = "Gene Id") %>%
  filter(GeneID != "None") %>%
  distinct(Gene_Symbol, .keep_all = TRUE)

write_csv(annotation, "gene_description.csv")
```

## Labels

We will load the header row of each tissue type to get the sample ids present. Then we will find the common samples among them and use only those for the analysis.
```{r message=FALSE, warning=FALSE}
tissue1 <- read_delim("../../../../HBTRC_data/GN326_MeanDataAnnotated_rev081815.txt", "\t", escape_double = FALSE, trim_ws = TRUE, skip = 32, n_max = 0) %>% select(starts_with("HB")) %>% names()
tissue2 <- read_delim("../../../../HBTRC_data/GN327_MeanDataAnnotated_rev081815.txt", "\t", escape_double = FALSE, trim_ws = TRUE, skip = 32, n_max = 0) %>% select(starts_with("HB")) %>% names()
tissue3 <- read_delim("../../../../HBTRC_data/GN328_MeanDataAnnotated_rev081815.txt", "\t", escape_double = FALSE, trim_ws = TRUE, skip = 32, n_max = 0) %>% select(starts_with("HB")) %>% names()

common_samples <- intersect(intersect(tissue1, tissue2), tissue3)
full_set <- union(union(tissue1, tissue2), tissue3)
```

Below is code to get the disease labels, i.e., whether the sample belongs to one among the class of Normal, Huntington's, or Alzheimer's.
```{r}
#Below is the format for getting label infromation from the sample id
pattern <- "^\\w{2}\\_\\w{3}\\_(\\w*)$"

#Now to get samples' labels using that format
labels <- sub(pattern, "\\1", common_samples)
labels <- recode(labels, "N"="Normal", "AD"="Alzheimer's", "HD"="Huntington's")

# label_df <- tibble(Samples=common_samples, Labels=labels)
```

We, however, in this bit want more relevant information with respect to just Alzheimer's. So, I will load in some meta-data with restricted access. It is still not personally recognizable, but it cannot be made public. No attempt at personally identifying has been made.
```{r message=FALSE, warning=FALSE}
meta <- read_csv("../../../../HBTRC_data/HBTRC_trait_data5_TID.txt")
matcher <- read_csv("../../../../HBTRC_data/patient_id_to_biosample_id.csv")

meta_data <- left_join(matcher, meta, by = c("Biosample Name")) %>%
                 filter(patient_id %in% common_samples)


table(meta_data$`AD case or control (combined)`)
table(meta_data$`Braak severity (from pathology reports)`) # 10 samples have NA values

labels_AD <- tibble(patient_id = meta_data$patient_id,
                    label = meta_data$`AD case or control (combined)`)

labels_braak <- tibble(patient_id = meta_data$patient_id,
                       label = meta_data$`Braak severity (from pathology reports)`) %>%
                  filter(!is.na(label))

labels_braak$label[labels_braak$label==1] = "1-2"
labels_braak$label[labels_braak$label==2] = "1-2"
labels_braak$label[labels_braak$label==3] = "3-4"
labels_braak$label[labels_braak$label==4] = "3-4"
labels_braak$label[labels_braak$label==5] = "5-6"
labels_braak$label[labels_braak$label==6] = "5-6"

table(labels_braak$label)
```


### Stratified Random Split
One for AD or not. One for Braak severity.
```{r}
# AD or Control
train.index <- createDataPartition(labels_AD$label, p = .8, list = FALSE)
train_AD <- labels_AD[ train.index,]
test_AD  <- labels_AD[-train.index,]

# Braak Severity
train.index <- createDataPartition(labels_braak$label, p = .8, list = FALSE)
train_braak <- labels_braak[ train.index,]
test_braak  <- labels_braak[-train.index,]
```

## Writing files

### The labels files
```{r}
# AD or not
labels_AD %>%
  rename(class = "label") %>%
  write.csv("labels_AD.csv", row.names = FALSE)

# Braak Severity
labels_braak %>%
  rename(class = "label") %>%
  write.csv("labels_braak.csv", row.names = FALSE)
```

## The train-test partition file
```{r}
# AD or not
fileConn<-file("trte_partition_AD.txt")
writeLines(c("training patient_id", paste(unlist(train_AD["patient_id"]), collapse = ","), "testing patient_id", paste(unlist(test_AD["patient_id"]), collapse = ",")), fileConn)
close(fileConn)

# Braak Severity
fileConn<-file("trte_partition_braak.txt")
writeLines(c("training patient_id", paste(unlist(train_braak["patient_id"]), collapse = ","), "testing patient_id", paste(unlist(test_braak["patient_id"]), collapse = ",")), fileConn)
close(fileConn)
```


