---
title: "METABRIC Gene Expression"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Importing all the requisite packages first.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(readr)
library(broom)
library(caret)
library(glmnet)
library(glmnetUtils)
library(foreach)
```

Some requisite functions.
```{r}
# To transpose
transpose_df <- function(df) {
  t_df <- data.table::transpose(df)
  colnames(t_df) <- rownames(df)
  rownames(t_df) <- colnames(df)
  t_df <- t_df %>%
    tibble::rownames_to_column(.data = .) %>%
    tibble::as_tibble(.)
  return(t_df)
}

# To perform ANOVA
run_anova <- function(train_data, train_labels, num_cpus=40) {
  num_cpus <- 40
  cl <- makeCluster(num_cpus)
  doParallel::registerDoParallel(cl)
  anova_summary <- foreach(i = 2:ncol(train_data)) %dopar% {
    column <- names(train_data[i]) 
    avz <- broom:: tidy(aov(train_data[,i][[1]] ~ unlist(train_labels$cancer_subtype))) # Each ANOVA test iterating through each column
    return(c(gene=column, f_value=avz$statistic[[1]], p_value=avz$p.value[[1]]))
  }
  stopCluster(cl)
  anova_summary <- as.data.frame(do.call(rbind, anova_summary), stringsAsFactors = FALSE)
  anova_summary <- transform(anova_summary, p_value = as.numeric(p_value), f_value = as.numeric(f_value))
  return(anova_summary)
}

# A greedy approach to selecting the top non-correlated features
greedySelect <- function(dframe, featureList, corrThreshold = 0.8, max_features = 1000){
  selected = c(featureList[1])
    for (i in c(2:length(featureList))) {
      tempList <- c(cor(dframe[featureList[i]], dframe[selected[1]])) 
      for (s in c(1:length(selected))){
        temp <- cor(dframe[featureList[i]], dframe[selected[s]])
        tempList <- append(tempList, temp)
        if (abs(temp) > corrThreshold) {break}
        }
      if (all( abs(tempList) < corrThreshold)) {
        selected <- append(selected, featureList[i])
      }
      if (length(selected) == max_features) {break}
    }
  return(selected)
  }
```

# Importing data

```{r}
colnames <- read_delim("../../../metabric/EGAF00000102986/discovery_ExpressionMatrix.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE, n_max = 1, col_names = FALSE)
colnames <- c("genes", as.character(colnames[1,]))

discovery <- read_delim("../../../metabric/EGAF00000102986/discovery_ExpressionMatrix.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE, col_names = colnames, skip = 1)

colnames <- read_delim("../../../metabric/EGAF00000102987/validation_ExpressionMatrix.txt", 
    " ", escape_double = FALSE, trim_ws = TRUE, n_max = 1, col_names = FALSE)
colnames <- c("genes", as.character(colnames[1,]))

validation <- read_delim("../../../metabric/EGAF00000102987/validation_ExpressionMatrix.txt", 
    " ", escape_double = FALSE, trim_ws = TRUE, col_names = colnames, skip = 1)

labels_discovery <- read_delim("clinical_data_discovery.txt", "\t") %>% 
                      dplyr::select(patient_id = METABRIC_ID, cancer_subtype = Pam50Subtype)
labels_validation <- read_delim("clinical_data_validation.txt", "\t") %>% 
                      dplyr::select(patient_id = METABRIC_ID, cancer_subtype = Pam50Subtype)
rm(colnames)
```


Converting the Illumina Gene IDs to Gene Symbols
```{r}
# BiocManager::install("illuminaHumanv3.db")
library(illuminaHumanv3.db)
genes <- discovery$genes
genes <- data.frame(Gene=unlist(mget(x = genes, envir = illuminaHumanv3ENTREZREANNOTATED)))

discovery$genes <- genes$Gene

genes <- validation$genes
genes <- data.frame(Gene=unlist(mget(x = genes, envir = illuminaHumanv3ENTREZREANNOTATED)))

validation$genes <- genes$Gene
rm(genes)
```
Removing NAs and summarising over duplicates
```{r}
discovery <- discovery[complete.cases(discovery), ] %>%
      group_by(genes) %>%
      summarize(across(everything(), list(mean))) #%>%
#      remove_rownames() %>%
#      column_to_rownames("genes")

validation <- validation[complete.cases(validation), ] %>%
      group_by(genes) %>%
      summarize(across(everything(), list(mean))) #%>%
#      remove_rownames() %>%
#      column_to_rownames("genes")

```

Reading in the requisite genes from the TCGA selected genes
```{r}
brca_genes <- read_csv("../TCGA BRCA/mrna_top1000.csv", n_max = 1, col_names = FALSE)

brca_genes <- as.character(brca_genes[,2:1001])
brca_genes <- unlist(str_split(brca_genes, "\\|"))[ c(FALSE, TRUE) ]

# metabric_genes <- as.character(discovery$genes)
# length(intersect(brca_genes, metabric_genes))
# length(setdiff(brca_genes, metabric_genes))
# setdiff(brca_genes, metabric_genes)
```

Selecting only those features selected for brca
```{r}
metabric_d <- discovery %>%
  filter(genes %in% brca_genes) %>%
  remove_rownames() %>%
  column_to_rownames("genes")

metabric_v <- validation %>%
  filter(genes %in% brca_genes) %>%
  remove_rownames() %>%
  column_to_rownames("genes")

setdiff(metabric_d$genes, metabric_v$genes)
```

Transposing and saving to files
```{r}
transpose_df(metabric_d) %>%
  dplyr::rename(patient_id = "rowname") %>%
  write.csv("metabric_discovery.csv", row.names = FALSE)

transpose_df(metabric_v) %>%
  dplyr::rename(patient_id = "rowname") %>%
  write.csv("metabric_validation.csv", row.names = FALSE)
```

Preparing for ANOVA
```{r}
metabric_d <- discovery %>%
  remove_rownames() %>%
  column_to_rownames("genes") %>%
  transpose_df() %>%
  dplyr::rename(patient_id = rowname)

metabric_d$patient_id <- unlist(str_split(metabric_d$patient_id, "\\_"))[c(TRUE, FALSE)]

metabric_d <- metabric_d[match(labels_discovery$patient_id, metabric_d$patient_id), ]
  
metabric_v <- validation %>%
  remove_rownames() %>%
  column_to_rownames("genes") %>%
  transpose_df() %>%
  dplyr::rename(patient_id = rowname)

metabric_v$patient_id <- unlist(str_split(metabric_v$patient_id, "\\_"))[c(TRUE, FALSE)]

metabric_v <- metabric_v[match(labels_validation$patient_id, metabric_v$patient_id), ]


identical(metabric_d$patient_id, labels_discovery$patient_id)
identical(metabric_v$patient_id, labels_validation$patient_id)
```

ANOVA
```{r}
anova_summary <- run_anova(metabric_d, labels_discovery, num_cpus=40)

discovery_sorted <- anova_summary %>% 
                  filter(p_value<=0.0001/nrow(anova_summary)) %>% 
                  filter(gene %in% names(metabric_v)) %>%
                  arrange(desc(f_value))

namelist <- c(discovery_sorted["gene"])
index <- match(namelist[[1]], names(metabric_d))
index <- prepend(index,1)     #To add the column with the patient_ids too to the output
metabric_d_anova <- metabric_d[,index]

#write.csv(metabric_d_anova, "metabric_d_anova.csv", row.names = FALSE)
```

Selecting top 1000 features
```{r}
metabric_best1000 <- greedySelect(metabric_d_anova, discovery_sorted["gene"][[1]], corrThreshold = 0.7)

index <- match(metabric_best1000, names(metabric_d_anova))
index <- prepend(index,1)     #To add the column with the patient_ids too to the output
metabric_d_top1000 <- metabric_d_anova[,index]
write.csv(metabric_d_top1000, "metabric_d_top1000.csv", row.names = FALSE)

index <- match(metabric_best1000, names(metabric_v))
index <- prepend(index,1)     #To add the column with the patient_ids too to the output
metabric_v_top1000 <- metabric_v[,index]
write.csv(metabric_v_top1000, "metabric_v_top1000.csv", row.names = FALSE)
```

