---
title: "Imputation Plots"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Imputation plots and selecting the best imputation methods

The Omics_Imputation.ipynb notebook in the folder was used to test a variety of methods and various parameters on a small set of missing omics, in order to find the best paramaters to the imputation methods. Simple readymade methods were tried.

The imputation.py file in the folder is used to run the methods with selected parameters on all combination of missing omics. It generates a csv file containing spearman correlation values for each missing omics combination, for each  methods, for each feature.

This file is used to generate a box-plot for the spearman correlation values, so as to determine the best imputation method to use.

Also, initially, RMSE and Stadard Deviation (a proxy for Variance) was used in place of Spearman Correlation for evaluating the imputation methods. The plots/csv files for the same can be found in "/logs/202104__/".


## Loading the data

```{r}
spearman <- read_csv("spearman.csv")
summary(spearman)
head(spearman)
```

## Processing the data

Adding a column stating what omics type any given feature belongs to.
```{r}
meth <- read_csv("../R/TCGA BRCA/meth_top1000.csv", n_max = 1)
mrna <- read_csv("../R/TCGA BRCA/mrna_top1000.csv", n_max = 1)
mirna <- read_csv("../R/TCGA BRCA/mirna_anova.csv", n_max = 1)

spearman$datatype <- NA
spearman$datatype[spearman$feature %in% names(meth)] = "meth"
spearman$datatype[spearman$feature %in% names(mrna)] = "mrna"
spearman$datatype[spearman$feature %in% names(mirna)] = "mirna"

head(spearman)
```

Removing the rows that represent features that were not missing, which consequently have zeros (in place of NaN) in the "Spearman" column.
```{r}
spearman <- spearman %>%
  rowwise() %>%
  filter(grepl(datatype, missing_values))

head(spearman)
```

## Box-plot

```{r}
spearman %>%
  ggplot(aes(y = Spearman, x = missing_values, fill = method)) +
  geom_boxplot()
```

