---
title: "Imputation Plots"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Imputation plots and selecting the best imputation methods

The Omics_Imputation.ipynb notebook in the folder was used to test a variety of methods and various parameters on a small set of missing omics, in order to find the best paramaters to the imputation methods. Simple readymade methods were tried.

The imputation.py file in the folder is used to run the methods with selected parameters on all combination of missing omics. It generates a csv file containing spearman correlation values for each missing omics combination, for each  methods, for each feature.

This file is used to generate a box-plot for the spearman correlation values, so as to determine the best imputation method to use.

Also, initially, RMSE and Stadard Deviation (a proxy for Variance) was used in place of Spearman Correlation for evaluating the imputation methods. The plots/csv files for the same can be found in "/logs/202104__/".


## Loading the data

```{r}
spearman <- read_csv("spearman.csv")
spearman2 <- read_csv("spearman2.csv")

spearman <- rbind(spearman, spearman2)
summary(spearman)
head(spearman)
```

## Processing the data

Adding a column stating what omics type any given feature belongs to.
```{r}
meth <- read_csv("../R/TCGA BRCA/meth_top1000.csv", n_max = 1)
mrna <- read_csv("../R/TCGA BRCA/mrna_top1000.csv", n_max = 1)
mirna <- read_csv("../R/TCGA BRCA/mirna_anova.csv", n_max = 1)

spearman$datatype <- NA
spearman$datatype[spearman$feature %in% names(meth)] = "meth"
spearman$datatype[spearman$feature %in% names(mrna)] = "mrna"
spearman$datatype[spearman$feature %in% names(mirna)] = "mirna"

head(spearman)
```

Removing the rows that represent features that were not missing, which consequently have zeros (in place of NaN) in the "Spearman" column.
```{r}
spearman <- spearman %>%
  rowwise() %>%
  filter(grepl(datatype, missing_omics))

spearman$method <- fct_relevel(spearman$method, c("random", "mean", "median", "enet", "knn_iter", "rf", "knn1", "knn25", "knn50", "knn75", "knn100", "knn200"))
spearman$missing_omics <- fct_relevel(spearman$missing_omics, c("mirna", "mrna", "meth", "mirna meth", "meth mrna", "mrna mirna"))

head(spearman)
```

## Box-plot

```{r}
spearman %>%
  ggplot(aes(y = Spearman, x = missing_omics, fill = method)) +
  geom_boxplot()+
  labs(x="Missing Omics", y="Correlation Coefficient (Spearman)", fill="Imputation method",
       title="Distribution of correlation with true values for imputed omics")+
  scale_fill_brewer(palette = "Paired", labels = c("Random", "Mean", "Median", "Elastic Net", "KNN (Iterative)", "Random Forest", "KNN (n=1)", "KNN (n=25)", "KNN (n=50)", "KNN (n=75)", "KNN (n=100)", "KNN (n=200)"))+
  scale_y_continuous(breaks = round(seq(-1, 1, by = 0.1),1))+
  scale_x_discrete(labels = c("miRNA", "mRNA", "DNA Methylation", "miRNA & Methylation", "Methylation & mRNA", "mRNA & miRNA"))
```


As can be inferred from the figure, KNN is the best choice to perform imputation with. 

KNN (Iterative) was performed with n=75. But if fails disproportionately badly on the "Methylation and mRNA" set, where the omics present was only miRNA. Similarly, KNN (n=25) performs particularly well for the same set.

Among the KNNs, averaging over 50 neighbours appears to be the best choice, while 25 neighbours is sometimes better. I believe that is so because in the training set, the smallest class contained less than 50 samples. Increasing "n" beyond that would likely result in including too many samples from other classes in the averaging process and thus reduce correlation and usefulness of the imputation in classification downstream.

This also implies that in case one trains the imputer on data sets containing more samples in the smallest class, one can increase "n" likewise for best results.
