---
title: "Data Split and Label Recovery"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

In this notebook, I will get the labels from the data and then format them as needed and also split the data into training and testing splits.

```{r}
library(tidyverse)
library(data.table)
library(readr)
library(broom)
library(caret)
```
## Loading data
```{r}
data <- read_delim("/data/users/bs16b001/HBTRC_data/GN326_MeanDataAnnotated_rev081815.txt", "\t", escape_double = FALSE, trim_ws = TRUE, skip = 32)
```


```{r}
# Getting the Gene Symbol and description and storing them as an annotation file
annotation <- data %>%
  select("Gene Symbol", "Gene Id", Description) %>%
  rename(Gene_Symbol = "Gene Symbol", GeneID = "Gene Id") %>%
  filter(GeneID != "None") %>%
  distinct(Gene_Symbol, .keep_all = TRUE)

write_csv(annotation, "gene_description.csv")
```
## Labels

We will load the header row of each tissue type to get the sample ids present. Then we will find the common samples among them and use only those for the analysis.
```{r}
tissue1 <- read_delim("/data/users/bs16b001/HBTRC_data/GN326_MeanDataAnnotated_rev081815.txt", "\t", escape_double = FALSE, trim_ws = TRUE, skip = 32, n_max = 0) %>% select(starts_with("HB")) %>% names()
tissue2 <- read_delim("/data/users/bs16b001/HBTRC_data/GN327_MeanDataAnnotated_rev081815.txt", "\t", escape_double = FALSE, trim_ws = TRUE, skip = 32, n_max = 0) %>% select(starts_with("HB")) %>% names()
tissue3 <- read_delim("/data/users/bs16b001/HBTRC_data/GN328_MeanDataAnnotated_rev081815.txt", "\t", escape_double = FALSE, trim_ws = TRUE, skip = 32, n_max = 0) %>% select(starts_with("HB")) %>% names()

common_samples <- intersect(intersect(tissue1, tissue2), tissue3)
```


```{r}
#Below is the format for getting label infromation from the sample id
pattern <- "^\\w{2}\\_\\w{3}\\_(\\w*)$"

#Now to get samples' labels using that format
labels <- sub(pattern, "\\1", common_samples)
labels <- recode(labels, "N"="Normal", "AD"="Alzheimer's", "HD"="Huntington's")

label_df <- tibble(Samples=common_samples, Labels=labels)
```

### Stratified Random Split
```{r}
train.index <- createDataPartition(label_df$Labels, p = .8, list = FALSE)
train <- label_df[ train.index,]
test  <- label_df[-train.index,]
```

## Writing files
### The disease_class file
```{r}
label_df %>%
  rename(patient_id = "Samples", class = "Labels") %>%
  write.csv("disease_class.csv", row.names = FALSE)
```

## The train-test partition file
```{r}
fileConn<-file("trte_partition.txt")
writeLines(c("training patient_id", paste(unlist(train["Samples"]), collapse = ","), "testing patient_id", paste(unlist(test["Samples"]), collapse = ",")), fileConn)
close(fileConn)
```

```{r}
labels$Category <- ifelse(labels$patient_id %in% train_idx, "Train", "Test")

labels %>%
  ggplot(aes(x=class, group=Category, fill=Category))+
  geom_bar(position="dodge")+
  labs(x="Disease class", y="Number of samples", title = "Sample distribution")

(Number_of_train_samples = length(train_idx))
(Number_of_test_samples = nrow(labels) - length(train_idx))
(Total_samples = nrow(labels))
```

